---
title: "immune cell atlas"
output: pdf_document
---

```{r load required packages, message=FALSE, warning=FALSE}
library(Seurat)
library(stats)
library(viridis)
library(ggplot2)
library(tidyverse)
```

### You will need to replace file paths with your own to reproduce results. See github page for data download instructions. 

```{r load fertile and infertile objects }
fertile <- readRDS("/Volumes/LaCie/seurat_objects/shiny_objects/fertile_atlas.rds")
infertile <- readRDS("/Volumes/LaCie/seurat_objects/shiny_objects/infertile_atlas.rds")
```

```{r subset immune cells from fertile and infertile objects}
fertile_immune <- subset(x = fertile, idents = 12)
infertile_immune <- subset(old, idents = c(1,3,4,5,6,7,10,11,12,13,16,17,18,19))
```

```{r integrate immune cells to generate immune atlas}
immune.anchors <- FindIntegrationAnchors(object.list = list(fertile_immune, infertile_immune), reduction = "cca", dims = 1:10)
immune <- IntegrateData(anchorset = immune.anchors, dims = 1:15)
DefaultAssay(immune) <- "integrated"
immune <- ScaleData(immune, assay = "integrated")
immune <- RunPCA(immune, features = VariableFeatures(object = immune))
immune <- RunUMAP(immune, reduction = 'pca', dims = 1:10)
immune <- FindNeighbors(immune, reduction = 'pca', dims = 1:10)
immune <- FindClusters(immune, resolution = 0.5)
```

```{r add fertility status metadata}
meta <- immune@meta.data
meta$cell.id <- row.names(meta)
meta[meta$age == '5 mo.', "fertility"] <- "fertile"
meta[meta$age == '12 mo.', "fertility"] <- "fertile"
meta[meta$age == '20 mo.', "fertility"] <- "fertile"
meta[meta$age == '22 mo.', "fertility"] <- "fertile"
meta[meta$age == '27 mo.', "fertility"] <- "infertile"
fertility <- as.vector(meta$fertility)
immune <- AddMetaData(immune, fertility, 'fertility')
```

# Cell types were assigned using marker genes and stored as a metadata column in the object. 
# The following code will reproduce the panels in Figures 3 and Supplemental 3.
# Objects with integrated cell type metadata are set to load here if one wishes to skip object generation and manual cell type assignment.

```{r immune cell atlas with metadata included}
immune <- readRDS("/Volumes/LaCie/seurat_objects/shiny_objects/immune_atlas.rds")
```


```{r Figure 3 Panel A}
my_levels <- c("T cells 1", "T cells 2", "T cells 3", "T cells 4",
               "T cells 5", "Treg", "macrophages 1",
               "macrophages 2", "macrophages 3", "macrophages 4",
               "leukocytes", "B cells", "neutrophils", "NK", "unknown immune")
immune@meta.data$cell.type <- factor(immune@meta.data$cell.type, levels = my_levels)


DimPlot(immune, group.by = 'cell.type', label = T, cols = c("#B0E0E6", "#4169E1", "#40E0D0", 
                                                            "#1E90FF", "#00008B","#87CEFA", 
                                                            "#FF69B4", "#FFC0CB", "#8B0000", 
                                                            "#DC143C", "#C8A2C8", "#98FB98", 
                                                            "#800080", "#2E8B57", "#CD5C5C"), pt.size = 1.5) + NoLegend() + NoAxes() + ggtitle(" ")
                                               
```

```{r Figure 3 panel B, message=FALSE, warning=FALSE}
DefaultAssay(immune) <- "RNA"
marker_list <- c("coro1a", "ptprc","lck", "gata3", "runx3", "zap70", 
                  "foxp3a", "cd4-1", 
                  "nkl.2", "pax5", 
                  "mpeg1.1", "csf1ra", "lgmn",
                  "mpx", "lyz",
                  "defbl1", "apodb")

my_levels_r <- c("unknown immune", "leukocytes","neutrophils", 
                 "macrophages 4", "macrophages 3", "macrophages 2",
                 "macrophages 1", "B cells", "NK", "Treg", "T cells 5",
                 "T cells 4", "T cells 3", "T cells 2", "T cells 1")
immune@meta.data$cell.type <- factor(immune@meta.data$cell.type, levels = my_levels_r)

DotPlot(immune, features = marker_list, scale = T,group.by = 'cell.type') + geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) + scale_colour_viridis(option="rocket", direction = -1) + guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white"))) + theme(axis.text.x = element_text(angle =45,hjust = 1)) + theme(legend.title = element_text( size = 8))
```
```{r Figure 3 panel C}
meta <- immune@meta.data
meta$cell.id <- row.names(meta)
fertile_immune_cells <- select(meta, c("cell.id", "fertility"))
fertile_immune_cells <- subset(fertile_immune_cells, fertility == 'fertile')
fertile_immune_cells <- as.vector(fertile_immune_cells$cell.id)
DimPlot(immune, cells.highlight = fertile_immune_cells, 
        cols.highlight = c('darkmagenta'), pt.size = 1.5) + NoLegend() + NoAxes()
```

# To generate the bar graph for panel D, the CellSelector function from Seurat was used to isolate somatic cell types from the fertile atlas. 

```{r selecting Leydig cells from fertile atlas}
# subset somatic cells from fertile atlas
fertile_soma <- subset(fertile, idents = c(11,12,13))

# first identify Leydig cell location with FeaturePlot()
FeaturePlot(fertile_soma, "rna_star")

plot <- DimPlot(object = fertile_soma)
# Follow instructions in the terminal to select points
Leydig <- CellSelector(plot = plot)
# cells are now stored in the character "Leydig"
Leydig
# the length of this character is the total amount of Leydig cells in the fertile atlas
```

```{r selecting macrophages from fertile atlas}
FeaturePlot(fertile_soma, "rna_csf1ra")
plot <- DimPlot(object = fertile_soma)
# Follow instructions in the terminal to select points
fertile_macrophages <- CellSelector(plot = plot)
# cells are now stored in the character "fertile_macrophages"
fertile_macrophages

# add new metadata column to differentiate immune cells types
meta <- fertile_soma@meta.data
meta$cell.id <- row.names(meta)
meta[meta$cell.id %in% fertile_macrophages, "cell.cat"] <- "macrophages"
cell.cat <- as.vector(meta$cell.cat)
# this now identifies macrophages by their cell "ID" which is generated by 10X
fertile_soma <- AddMetaData(fertile_soma, cell.cat, "cell.category")

# somatic cell types were grouped broadly by category and renamed in metadata 
fertile_soma@meta.data$cell.type[fertile_soma@meta.data$cell.type == 'Sertoli',] <- "SLV"
fertile_soma@meta.data$cell.type[fertile_soma@meta.data$cell.type == 'Leydig',] <- "SLV"
fertile_soma@meta.data$cell.type[fertile_soma@meta.data$cell.type == 'vSMC',] <- "SLV"
```

```{r isolate numbers for each cell type from each age in the fertile atlas}
# separate fertile soma by age
meta <- fertile_soma@meta.data
meta$cell.id <- row.names(meta)
meta <- select(meta, c(cell.id, age, cell.type, cell.category))
mo5_soma_meta <- subset(meta, meta$age == '5 mo.')
yr1_soma_meta <- subset(meta, meta$age == '12 mo.')
yr1.6_soma_meta <- subset(meta, meta$age == '20 mo.')
yr1.8_soma_meta <- subset(meta, meta$age == '22 mo.')

# numbers for each immune cell category from each age are displayed
table(mo5_soma_meta$cell.category)
table(yr1_soma_meta$cell.category)
table(yr1.6_soma_meta$cell.category)
table(yr1.8_soma_meta$cell.category)
```

```{r isolate numbers for somatic cell types in the infertile atlas}
# macrophages example
table(infertile@meta.data$cell.type)
2147 + 1508 + 1039  
# 4694 macrophages
```

```{r build dataframe for bargraph from cell numbers}
# the cell percent is determined by dividing the cell count by the total number of cells in the individual age's object
cell.cats <- c("SLV", "macrophages", "lymphocytes", "other immune", "blood")
counts <- c(200, 31, 56, 35, 22)
percents <- counts/12650*100
age <- c("5 mo.", "5 mo.", "5 mo.", "5 mo.","5 mo.")
mo5_df <- data.frame(age, cell.cats, counts, percents)


counts <- c(138, 17, 62, 18, 53)
percents <- counts/14084*100
age <- c("12 mo.", "12 mo.", "12 mo.", "12 mo.","12 mo.")
mo12_df <- data.frame(age, cell.cats, counts, percents)


counts <- c(28, 10, 14, 15, 4)
percents <- counts/2636*100
age <- c("20 mo.", "20 mo.", "20 mo.", "20 mo.","20 mo.")
mo20_df <- data.frame(age, cell.cats, counts, percents)


counts <- c(60, 27, 47, 23, 5)
percents <- counts/3289*100
age <- c("22 mo.", "22 mo.", "22 mo.", "22 mo.","22 mo.")
mo22_df <- data.frame(age, cell.cats, counts, percents)


counts <- c(1381, 4694, 8804, 605, 1222)
percents <- counts/32659*100
age <- c("27 mo.", "27 mo.", "27 mo.", "27 mo.","27 mo.")
mo27_df <- data.frame(age, cell.cats, counts, percents)

# combine dataframes 
soma_df <- rbind(mo5_df, mo12_df, mo20_df, mo22_df, mo27_df)
soma_df$age <- factor(soma_df$age, levels = c("5 mo.", "12 mo.", "20 mo.", "22 mo.", "27 mo."))
```

```{r plot relative abundance}
ggplot(soma_df, aes(fill=age, x=factor(cell.cats, level=c("SLV", "macrophages", "lymphocytes", "other immune", "blood")), y=percents)) + 
  geom_bar(position = 'dodge', stat = 'identity', color = 'black') + theme_classic() + 
  ylab("normalized percent") + xlab("cell type") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=.5)) + scale_fill_manual(values=c('#F0F8FF', '#ADD8E6', '#87CEEB', '#4682B4', '#204a87')) + ggtitle("relative abundance of somatic cell types across fertility")
```
```{r Supplemental Figure 3 Panel A}
DimPlot(immune, split.by = 'age', cols = c("#B0E0E6", "#4169E1", "#FF69B4", "#FFC0CB","#2E8B57", "#87CEFA",
                 "#40E0D0", "#8B0000","#DC143C", "#800080", "#98FB98","#C8A2C8",
                 "#1E90FF", "#CD5C5C", "#00008B")) + NoLegend() + NoAxes()
```


```{r Supplemental Figure 3 Panel B}
my_levels <- c("T cells 1", "T cells 2", "T cells 3", "T cells 4",
               "T cells 5", "Treg", "macrophages 1",
               "macrophages 2", "macrophages 3", "macrophages 4",
               "leukocytes", "B cells", "neutrophils", "NK", "unknown immune")
immune@meta.data$cell.type <- factor(immune@meta.data$cell.type, levels = my_levels)
marker_list <- c("coro1a", "ptprc","lck", "gata3", "runx3", "zap70", 
                  "foxp3a", "cd4-1", 
                  "nkl.2", "pax5", 
                  "mpeg1.1", "csf1ra", "lgmn",
                  "mpx", "lyz",
                  "defbl1", "apodb")
VlnPlot(immune, marker_list, group.by = 'cell.type', stack = T, flip = T) + NoLegend()
```
```{r Supplemental Figure 3 Panel C-F, message=FALSE, warning=FALSE}
DefaultAssay(fertile) <- "RNA"
FeaturePlot(fertile, "anxa5b", pt.size = 1, order = F) + NoAxes() + ggtitle("anxa5b") + scale_colour_viridis(option = "rocket", direction = -1)
FeaturePlot(infertile, "anxa5b", pt.size = 1, order = F) + NoAxes() + ggtitle("anxa5b") + scale_colour_viridis(option = "rocket", direction = -1)
VlnPlot(infertile, "anxa5b", pt.size = 0, group.by = 'cell.type') + ggtitle("anxa5b") + NoLegend()
VlnPlot(infertile, "anxa5b", pt.size = 0, group.by = 'cell.type') + ggtitle("anxa5b") + NoLegend()
```
```{r Supplemental Figure 3 G, message=FALSE, warning=FALSE}
immune <- ScaleData(immune, assay = 'RNA')
Tcells1 <- c("ccl34b.4", "ccl38a.4", "crema", "tnfrsf9a", "traf1")
Tcells2 <- c("sat1a.1", "ccl36.1", "ca2", "il2rb", "fosab")
macrophages1 <- c("lgals2a", "lygl1", "cd74a", "c1qc", "c1qb")
macrophages2 <- c("ccl35.1", "hspb1", "marcksl1b", "timp2b", "fthl27")
NK <- c("nkl.2", "il1fma", "ccl36.1", "ccl38.6", "slc41a1")
Treg <- c("foxp3a", "il1rl1", "cd4-1", "cd28l", "cd28")
Tcells3 <- c("zp2.5", "setbp1", "crp2", "org", "hbba1")
macrophages3 <- c("h1m", "zp3", "cldng", "tmsb5", "ccnb1")
macrophages4 <- c("grn2", "lgmn", "grn1", "ctsd", "gpnmb")
neutrophils <- c("lgals2b", "krt91", "krt92", "krt4", "cldnh")
Bcells <- c("igic1s1", "sftpba", "p2rx3a", "ctsbb", "ccl35.1")
leukocytes <- c("spink2.2", "c3a.4", "itln1", "cfh", "c3a.5")
Tcells4 <- c("zgc:110216", "hist1h2a6", "zgc:113984.1", "hist2h3c", "hist1h4l")
unknown <- c("apodb", "defbl1", "lypc", "apoda.1", "alx4a")
Tcells5 <- c("il4", "il13", "il11b", "csf3a", "gata3")

features_list <- c(Tcells1, Tcells2, Tcells3, Tcells4, Tcells5, Treg, NK,
                   Bcells, macrophages1, macrophages2, macrophages3, macrophages4, 
                   neutrophils, leukocytes, unknown)

#my_levels <- c("T cells 1", "T cells 2", "T cells 3", "T cells 4", "T cells 5", 
               #"Treg", "NK", "B cells", "macrophages 1", "macrophages 2", 
               #"macrophages 3", "macrophages 4", "neutrophils", "leukocytes", "unknown immune")
#immune@meta.data$cell_type <- factor(immune@meta.data$cell_type, levels = my_levels)

my_levels <- c(0, 1, 6, 12, 14, 5, 4, 10, 2, 3, 7, 8, 9, 11, 13)
immune@meta.data$seurat_clusters <- factor(immune@meta.data$seurat_clusters, levels = my_levels)
#immune <- ScaleData(immune, assay = "RNA")

DoHeatmap(immune, assay = 'RNA', slot = "scale.data", features = features_list, raster = F, lines.width = 20, label = F, size = 2, angle = 90, group.by = "seurat_clusters", group.colors = c("#B0E0E6", "#4169E1", "#40E0D0", 
                                                            "#1E90FF", "#00008B","#87CEFA", 
                                                            "#FF69B4", "#FFC0CB", "#8B0000", 
                                                            "#DC143C", "#C8A2C8", "#98FB98", 
                                                            "#800080", "#2E8B57", "#CD5C5C"),draw.lines = T) + theme(axis.text.y = element_text(size = 7)) + scale_fill_viridis_c(option="rocket", direction = 1)
```


